---
title: "Manual Annotation on GW9"
author: Yuji Okano
date: September 2, 2022
output:
  md_document:
    variant: markdown_github
---

# 05-3_manual_annot

## Objectives

1.  Manual Annotation via Seurat

### load data and make seurat object

```{r}
library(dplyr)
library(jsonlite)
library(Seurat)
library(patchwork)


source("../tools/spongy_panda/export_gdcmatrix.R")

data.id = "gw12"
data <- readRDS(paste0("../../data/gse165388_variablefeat/", data.id, "feat_seurat_pbj.rds"))

dim <- as.integer(read_json("../../out/gse165388/05_seurat_annot/dimensionality.json", simplifyVector = TRUE)[data.id])
```

### make difrectory to save outputs
```{r}
dir.name <- "../../data/gse165388_annot"

if (! dir.exists(dir.name)) {
  dir.create(dir.name)
}
```

## PCA
```{r}
all.genes <- rownames(data)
feat <- VariableFeatures(object = data)

data <- RunPCA(
  data,
  features = feat,
  verbose = FALSE,
  seed.use = 0,
  npcs=length(feat),
  approx=FALSE
)
```

## Clustering
```{r}
data <- FindNeighbors(data, dims = 1:dim)
data <- FindClusters(data, resolution = 0.8)
```

### Get cluster id
```{r}
clusters_id <- data.frame(Idents(data))
```

## Embedding into Manifolds
### UMAP
```{r}
data <- RunUMAP(data, dims = 1:dim)

DimPlot(data, reduction = "umap", label = T)
```

## Find DEG
```{r}
markers <- FindAllMarkers(data, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
n <- 10

degs <- markers %>%
  group_by(cluster) %>%
  slice_max(n = n, order_by = avg_log2FC) %>%
  data.frame()

write.csv(
  degs, 
  paste0(dir.name, "/", data.id, "_degs_top", n, ".csv"),
  quote = F
  )
```

### choose one key gene per cluster
```{r}
deg.list <- sapply(unique(degs$cluster), function(idx) degs[degs$cluster==idx, ]$gene[1])
```


## Visualization

```{r, fig.height=12, fig.width=24, dpi=80, out.width=24, out.height=12}
VlnPlot(data, features = deg.list)
```

- scatter plot

```{r, fig.height=24, fig.width=24, dpi=80, out.width=24, out.height=24}
FeaturePlot(data, features = deg.list, label=T)
```

- heatmap

```{r, fig.height=10, fig.width=30, dpi=80, out.width=30, out.height=10}
markers %>%
    group_by(cluster) %>%
    top_n(n = 3, wt = avg_log2FC) -> top
DoHeatmap(data, features = top$gene)
```

- After these procedure, clusters are **manually** annotated

```{r}
degs
```

### Assign Names
```{r}
new.label <- c(
  "Unclassifiable", "MGE", "LGE", "NPC", "NPC", "LGE",
  "MGE", "Unclassifiable", "NPC", "Excitatory", "Unclassifiable", "NPC",
  "CGE", "Endothelial", "RBC", "Microglia"
)
```

### Visualization
```{r}
names(new.label) <- levels(data)
data <- RenameIdents(data, new.label)
DimPlot(data, reduction = "umap", label = TRUE, pt.size = 0.5)
```

### Export Clusters 
```{r, include=FALSE}
annot <- data.frame(Idents(data))
colnames(annot) <- c("annot")
colnames(clusters_id) <- c("cluster_id")

annot <- data.frame(clusters_id, annot)
rownames(annot) <-  rownames(clusters_id)

write.table(
  annot,
  paste0(dir.name, "/", data.id, "_clusters.csv"),
  append = F,
  sep = ",", 
  row.names = T,
  quote = F
)
```


## Export Data
```{r}
saveRDS(Embeddings(data, reduction = "pca")[, 1:dim], file = paste0(dir.name, "/", data.id, "pca.rds"))
saveRDS(Embeddings(data, reduction = "umap"), file = paste0(dir.name, "/", data.id, "umap.rds"))
```