---
title: "Manual Anntation on GW9"
author: Yuji Okano
date: August 24, 2022
output:
  md_document:
    variant: markdown_github
---

# 05_manual_annot

## Objectives

1.  Manual Annotation via Seurat

### load data and make seurat object

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)

data <- readRDS("../../data/gse165388_processed/gw9_seuratobject.rds")
```

## Data scaling (for PCA)
```{r}
all.genes <- rownames(data)
data <- ScaleData(data, features = all.genes)
```

## PCA
```{r}
data <- RunPCA(
  data,
  features = all.genes,
  verbose = FALSE,
  seed.use = 0,
  npcs=length(all.genes),
  approx=FALSE
)
```


```{r}
plot(Stdev(data, reduction = "pca"))
```


```{r}
# Examine and visualize PCA results a few different ways
print(pbmc[["pca"]], dims = 1:5, nfeatures = 5)
```

### visualizaition
- loadings
```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

- scatter plot
```{r}
DimPlot(pbmc, reduction = "pca")
```


### determine n_dim
- Jack Straw plot
```{r}
# NOTE: This process can take a long time for big datasets, comment out for expediency. More
# approximate techniques such as those implemented in ElbowPlot() can be used to reduce
# computation time
pbmc <- JackStraw(pbmc, num.replicate = 100)
pbmc <- ScoreJackStraw(pbmc, dims = 1:20)
JackStrawPlot(pbmc, dims = 1:15)
```

- Elbow plot
```{r}
ElbowPlot(pbmc)
```

- **Note**: when the data is small enough, `fa.parallel` from `psych` is useful

## Clustering
```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```

- checking ids
```{r}
# Look at cluster IDs of the first 5 cells
head(Idents(pbmc), 5)
```

## Embedding into Manifolds
### UMAP
```{r}
# If you haven't installed UMAP, you can do so via reticulate::py_install(packages =
# 'umap-learn')
pbmc <- RunUMAP(pbmc, dims = 1:10)

# note that you can set `label = TRUE` or use the LabelClusters function to help label
# individual clusters
DimPlot(pbmc, reduction = "umap")
```

### sSNE
```{r}
# If you haven't installed UMAP, you can do so via reticulate::py_install(packages =
# 'umap-learn')
pbmc <- RunTSNE(pbmc, dims = 1:10)

# note that you can set `label = TRUE` or use the LabelClusters function to help label
# individual clusters
DimPlot(pbmc, reduction = "tsne")
```

## Export results so far

```{r}
saveRDS(pbmc, file = "../../data/tutorial/pbmc_tutorial.rds")
```

-   **Note**: .rds format can be used in python as well. use `pyreadr`

## Find DEG
- markers of cluster2

```{r}
# find all markers of cluster 2
cluster2.markers <- FindMarkers(pbmc, ident.1 = 2, min.pct = 0.25)
head(cluster2.markers, n = 5)
```

- markers of cluster5 compared to cluster 0-3

```{r}
# find all markers distinguishing cluster 5 from clusters 0 and 3
cluster5.markers <- FindMarkers(pbmc, ident.1 = 5, ident.2 = c(0, 3), min.pct = 0.25)
head(cluster5.markers, n = 5)
```

- iteratively compare all clusters

```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
```

- ROC can be a measure of differential expressions as explanatory variables

```{r}
cluster0.markers <- FindMarkers(pbmc, ident.1 = 0, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
cluster0.markers
```

## Visualization

```{r}
VlnPlot(pbmc, features = c("MS4A1", "CD79A", "GAPDH", "TAPBPL", "GADD45G", "POLR1A"))
```

- violing plot of rawcounts

```{r}
# you can plot raw counts as well
VlnPlot(pbmc, features = c("NKG7", "PF4", "POLR2A"), slot = "counts", log = TRUE)
```

- scatter plot

```{r}
FeaturePlot(pbmc, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "PPBP",
    "CD8A", "UGT8", "CD44"))
```

- heatmap

```{r}
pbmc.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```

- After these procedure, clusters are **manually** annotated

```{r}
new.cluster.ids <- c("Naive CD4 T", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
    "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

### Expot all results

```{r}
saveRDS(pbmc, file = "../../data/tutorial/pbmc_tutorial_all.rds")
```

## Cmds of RNotebook

-   *Insert Chunk*: `Opt+Cmd+I`
-   *Run*: `Shift+Cmd+Return`
-   *Preview*: `Shit+Cmd+K`
-   *Save*: `Cmd+S`
